Comparative Evaluation of Spark and Flink Stream Processing

### 2017-03-07 paper/paper.tex
[paper] updated the abstract and introduction

\end{abstract}
Recent years have witnessed a advances in the  development of Big Data frameworks like Apache Spark and Apache Flink. For instance, both frameworks support real-time
streaming processing. While which platform is the best still an open question. 
In this report, we provide a performance comparison of streaming data processing in Apache Spark and Apache Flink, by measuring different performance metrics, namely latency and throughput. In addition, we cover some key aspects of streaming data applications and how they are handled in both frameworks. The experiment is done using multiple streaming workloads over real-world datasets.
\section{Introduction}

\par In recent years, Big Data is considered as one of the most appealing topics in Information Technology, it has gained increasing attention of professionals in industry and academia, as a result of the fact that the amount of data we generate is increasing exponentially \cite{idc} that generated from social media platforms, mobile phones, IoT and finical transactions, etc. Moreover, recent years showed a rapid development of Big Data frameworks such as Apache Spark and Apache Flink that provide more and efficient data processing capabilities.
 \par The Big Data is a method to process not just only large datasets (volume), but also the data that arrives in high rate (velocity), and it comes in all kinds of data format (variety) (i.e., 3Vs of Big Data (Volume, Velocity, Variety)) \cite{svs}.

\par The processing models of Big Data are Batch processing and Stream processing, Batch processing involves handling  large, finite volumes of data (e.g., processing of historical transaction records), is more concerned with throughput. While the Stream processing  is a method to manage fast and continuously incoming data (real-time), process it as it arrives to the system (e.g., credit card fraud detection and network monitoring applications), where the low latency is required. The variety of data (structured and semi-structured) is included in the two processing models.

\par The are some performance evaluation of Apache Spark and Apache Flink  such as Yahoo Streaming Benchmark \cite{yahoo}. We believe that we still need to carry our evaluation for two reasons: \begin{enumerate*}[label=(\roman*)]
\item both frameworks are continuously improving in short time 
\item we aim to study the performance of some complex operations such as \texttt{groupByKey} \& \texttt{keyBy} in Apache Spark and Apache Flink, respectively.
\end{enumerate*}

\par The purpose of this report is to provide a comparative experimental evaluation of throughput, latency (i.e., performance) of stream processing in Apache Spark and Apache Flink. We use a datasets of airplanes trajectories provided in the context of Datacron project\footnote{\url{http://www.datacron-project.eu/}}, in order to simulate real-world use cases. Using multiple streaming data workloads. Also we try to cover general aspects of streaming processing tasks (later details in Section 3) and how they can be solved in each platform. Informally,  this work aims to help the developers to determine which platform to choose in production for different use cases.

\par The remainder of this report is organized as follows.
In Section 2, we present the main characteristics, APIs, and main data abstraction of Apache Spark and Flink. Section 3  presents the general aspects of streaming processing, the design and implementation of experimental workloads. Section 4
provides and compares the performance results of the different workloads. Last, Section 5 gives the overall conclusion.

>>>>

\end{abstract}
Recent years have witnessed advances in the  development of Big Data frameworks like Apache Spark and Apache Flink. For instance, both frameworks support real-time data 
stream processing. The question of which platform is superior to the other still open. 
In this report, we provide a performance comparison of streaming data processing in Apache Spark and Apache Flink by measuring different performance metrics (latency and throughput). In addition, we cover some key aspects of streaming data applications and how they are handled in the two frameworks. The experiment is done using multiple streaming workloads over real-world datasets.
\section{Introduction}

\par In recent years, Big Data is considered as one of the most appealing topics in Information Technology, it has gained increasing attention of professionals in industry and academia. As result of the fact that the amount of data we generate is increasing exponentially \cite{idc}.  Different sources such as  social media platforms, mobile phones, IoT and finical transactions, etc.  create large volumes of data.  Moreover, recent years have shown a rapid development of Big Data frameworks such as Apache Spark and Apache Flink that provide efficient data processing functionalities.
 \par Big Data is a term that does not only describe large datasets (volume), but also the data that arrives in high rate (velocity), which comes in all kinds of data format (variety) (i.e., 3Vs of Big Data (Volume, Velocity, Variety)) \cite{svs}.

\par The processing models of Big Data are Batch processing and Stream processing. Batch processing involves handling  large, finite volumes of data (e.g., processing of historical transaction records), and is more concerned with throughput. While Stream processing  is a method to manage fast and continuously incoming data (real-time), as it arrives to the system (e.g., credit card fraud detection and network monitoring applications). Where low latency is crucial for stream processing systems. Both processing models are capable of operating with structured or semi-structured data (i.e., variety).

\par There exist some performance evaluations of Apache Spark and Apache Flink  such as Yahoo Streaming Benchmark \cite{yahoo}. We believe that we still need to carry our evaluation for two reasons: \begin{enumerate*}[label=(\roman*)]
\item both frameworks are continuously improving in short time 
\item we aim to study the performance of some complex operations such as \texttt{groupByKey} \& \texttt{keyBy} in Apache Spark and Apache Flink, respectively.
\end{enumerate*}

\par The purpose of this report is to provide a comparative experimental evaluation of throughput and latency of stream processing in Apache Spark and Apache Flink. In order to simulate real-world use cases, we use datasets of airplane trajectories provided by the  Datacron project\footnote{\url{http://www.datacron-project.eu/}} to develop multiple streaming data processing workloads. Beside the performance comparison of stream processing of the two frameworks, we try to cover general aspects of stream processing tasks (see Section 3) and how they can be solved in each platform. Informally,  this work aims to help the developers to determine which platform to use for different use cases.

\par The remainder of this report is organized as follows.
In Section 2, we present the main characteristics, programming APIs, and main data abstraction of Apache Spark and Apache Flink. Section 3  presents the general aspects of streaming processing and the implementation of experimental workloads. Section 4
compares the performance results of the different workloads. Section 5 gives the overall conclusion.
